### 2012-09-12 paper_sean/conclusion.tex
lots of cuts, down to 13 pages.

% !TEX root =main.tex
HyperNEAT has proven to be an effective encoding in gait evolution which is able to find gait regularities without human intervention \cite{clune1,clune2,yos:clune}. 
Glette et al. \cite{glette} has shown that evolving gaits in simulation using a simple algorithm with a large number of evaluations could outperform gait evolution using a more sophisticated algorithm in hardware.  
This was possible because the simulator afford a much larger number of evaluations to be performed during the evolution process. 
In this study, we took the two -- a sophisticated algorithm (HyperNEAT) and a physical simulator -- to see how HyperNEAT would perform with a many evaluations in the evolution process. 
We showed that HyperNEAT, when used with a simulator, could evolve very impressive gaits. 
In simulation, HyperNEAT gaits outperformed Genetic Algorithm gaits \cite{glette} by 54.3\%. 
In the real life trials, HyperNEAT's best gait performed 5.4\% better than the GA gait. 
This study also showed that HyperNEAT will exploit the characteristics of the problem, such as ferquency and regularity, in order to maximize its fitness.
%Without frequency limiters, it would make the legs move really fast in order to increase fitness. 
%When that was no longer allowed, it pushed the servo positions to their boundaries by increasing the stride lengths as wide as possible. 
However, because of the physical limitations of the robot, these exploitations did not transfer well to reality. %could not be transferred well to reality.
Moreover, the inevitable faults with the simulator (it cannot model the robot or the real world perfectly) added to the difficulty in the transfer from simulation to reality. 
Thus, many of the gaits performed much worse than it did in simulation and had difficulty in achieving consistent results.

>>>>

% !TEX root =main.tex
HyperNEAT has proven to be an effective encoding in gait evolution with an ability to find gait regularities without human intervention \cite{clune1,clune2,yos:clune}. 
Glette et al. \cite{glette} has shown that evolving gaits in simulation using a simple algorithm with a large number of evaluations could outperform gait evolution using a more sophisticated algorithm in hardware thanks to the simulator affording a larger number of evaluations to be performed for evolution. 
In this study, we took the two -- HyperNEAT and a physical simulator -- and showed that HyperNEAT, when it could afford many evaluations, could evolve impressive gaits. 
In simulation, HyperNEAT gaits outperformed Genetic Algorithm gaits \cite{glette} by 54.3\% and by 5.4\% in the real life trials. 
This study also showed that HyperNEAT will exploit the characteristics of the problem, such as ferquency and regularity, in order to maximize its fitness.
%Without frequency limiters, it would make the legs move really fast in order to increase fitness. 
%When that was no longer allowed, it pushed the servo positions to their boundaries by increasing the stride lengths as wide as possible. 
However, because of the physical limitations of the robot, these exploitations did not transfer well to reality. %could not be transferred well to reality.
Moreover, the inevitable faults with the simulator (it cannot model the robot or the real world perfectly) added to the difficulty in the transfer from simulation to reality. 
Thus, many of the gaits performed much worse than it did in simulation and had difficulty in achieving consistent results.

### 2012-09-12 paper_sean/results.tex
lots of cuts, down to 13 pages.

% !TEX root =main.tex
HyperNEAT, used with a physical simulator, produced gaits that outperformed all previous gaits tried and recorded on the QuadraTot.\footnote{Videos of the gaits in simulation and on robot available at http://quadratot.yosinski.com}
The speed of the fastest gait recorded in this study was 14.5 cm/s, faster than the previously fastest gait published in Glette et al. \cite{glette}. 
Kyrre et al. \cite{glette} recorded 17.8 cm/s on their robot, but to make sure that there were no differences in robots, the same gait was also tested on the robot used for this study (which was the one used by Yosinski et al. \cite{yos:clune} and Shen et al. \cite{haocheng}. 
The same gait recorded an average speed of 12.951 cm/s and a maximum of 13.76 cm/s on our robot. We used these numbers in comparing results. 
The gait from this study outperformed the previous Kyrre et al's \cite{glette} gait by 5.4\% with much fewer evaluations performed per run, at 40000 evaluations per run as opposed to Kyrre et al's \cite{glette} 60000 evaluations per run.
It also outperformed the best gait evolved using HyperNEAT in hardware \cite{yos:clune} by 49.5\%, most likely due to the difference in the number of evaluations between the two studies -- 40000 vs 180. 
Another minor reason for superior performance may be that frequency limiter policy that we placed may have helped in producing gaits that were less susceptible to shutdown while Yosinski et al. \cite{yos:clune} did not have such things.
Although most gaits started off its evolution process by producing fast and high frequency gaits, with more generations, the gaits dramatically reduced its frequencies while continually increasing their fitness. The average frequency of the final gaits was 1.29 Hz per servo.

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=fit_gen.eps, width=8cm}
% \caption[ ]{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%.}
% \end{center}
% \end{figure}
\figp{aracna_notYetCleanedMultiPieceRobot_edit}{.6}{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%. \edit{wrong fig, regenerate as pdf and add.}}

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=freq_fitness.eps, width=8cm}
% \caption[ ]{Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot.}
% \end{center}
% \end{figure}
\figp{aracna_notYetCleanedMultiPieceRobot_edit}{.6}{Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot. \edit{wrong fig, regenerate as pdf and add.}}

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=servo_pos.eps, width=8cm}
% \caption[ ]{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al. \cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}
% \end{center}
% \end{figure}
\figgp{aracna_notYetCleanedMultiPieceRobot_edit}{.6}{neat_110115_211410_00000_002_filt_zoom}{.6}{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al. \cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem. \edit{wrong fig on top, regenerate top fig as pdf and add.}}

>>>>

% !TEX root =main.tex
HyperNEAT, used with a physical simulator, produced gaits that outperformed all previous gaits tried and recorded on the QuadraTot.\footnote{Videos of the gaits in simulation and on robot available at http://quadratot.yosinski.com}
The fastest gait published was from Kyrre et al. \cite{glette}, which recorded 17.8 cm/s on their robot. 
For consistency, we took that gait and ran it ten times on our robot, and found that on our robot, the gait had an average speed of 12.95 cm/s and a maximum speed of 13.76 cm/s.
The speed of the fastest gait recorded in this study was 14.5 cm/s, which is 5.4\% faster than Kyrre et al.'s \cite{glette} gait with 20000 less evaluations performed per run.
%Kyrre et al. \cite{glette} published 17.8 cm/s on their robot, but to make sure that there were no differences in robots, the same gait was also tested on our robot which then recorded an average speed of 12.951 cm/s and a maximum speed of 13.76 cm/s. 
%The gait from this study outperformed the previous Kyrre et al's \cite{glette} gait by 5.4\% with much fewer evaluations performed per run, at 40000 evaluations per run as opposed to Kyrre et al's \cite{glette} 60000 evaluations per run.
It also outperformed the best gait evolved using HyperNEAT in hardware \cite{yos:clune} by 49.5\%, most likely due to the difference in the number of evaluations between the two studies -- 40000 vs 180. 
Another minor reason for superior performance may be that frequency limiter policy that we placed may have helped in producing gaits less susceptible to shutdown while Yosinski et al. \cite{yos:clune} had no such implementation.
Although most gaits started its evolution process by producing fast and high frequency gaits, with more generations, the gaits dramatically reduced its frequencies while continually increasing their fitness. The average frequency of the final gaits was 1.29 Hz per servo.

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=fit_gen.eps, width=8cm}
% \caption[ ]{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%.}
% \end{center}
% \end{figure}
\figp{avg_fit_vs_gen}{.6}{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%.}% \edit{wrong fig, regenerate as pdf and add.}}

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=freq_fitness.eps, width=8cm}
% \caption[ ]{Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot.}
% \end{center}
% \end{figure}
\figp{avg_freq_vs_gen}{.6}{Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot.} %\edit{wrong fig, regenerate as pdf and add.}}

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=servo_pos.eps, width=8cm}
% \caption[ ]{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al. \cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}
% \end{center}
% \end{figure}
%\figgp{servo_plot_111}{.6}{neat_110115_211410_00000_002_filt_zoom}{.6}{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al. \cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}%\edit{wrong fig on top, regenerate top fig as pdf and add.}}
\figp{servo_plot_111}{.6}{A plot of servo positions for a gait produced in this study. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}%\edit{wrong fig on top, regenerate top fig as pdf and add.}}

### 2012-09-12 paper_sean/robot.tex
lots of cuts, down to 13 pages.

The Dynamixel AX-12A and AX-18A servos have a safety feature called \emph{Alarm Shutdown} that would turn off the servo if the servo's current, range, temperature, or torque was too high \cite{robotis}. 
The ones most relevant to this study were shutdowns due to high torque and range. 
As has been reported previously with QuadraTot \cite{yos:clune}, many of the gaits that were tested on the QuadraTot forced the servos to exert too much torque or would command out of range positions and cause the servos to shutdown. In order to address this problem and avoid collisions with its own body, a motion cropping system was implemented on the QuadraTot. 
In addition to placing limits on the range of motions for the inner joints, outer joints, and the hip joint of [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree] respectively, an extra cropping system named the smart cropping system was adopted from Shen et al. \cite{haocheng} to further prevent servo shutdowns from leg positions where the combination of the two angles interact to produce an extreme amount of torque. This smart cropping algorithm calculates a leg position by summing the servo position values from the inner and outer joints of a leg, \emph{\textbf{g}} (\emph{\textbf{g}} = $j_{inner}$ + $j_{outer}$) %JMC/SL - do they sum the position values, or the joint angles? Since all the previous extremes are expressed in angles, the writing makes me expect angles... / They do use the position values and I think that it would be best to be consistent with what Jason/Haocheng did. This is also the way g is calculated in the simulator and the python drivers as well and more accurately reflects what we actually did. 
and modifying the servo positions from ``fierce gaits'', which usually have relatively low \emph{\textbf{g}}s, to remap them onto the boundary of a quasi-trangular area in a 2-D space defined by $j_{inner}$ and $j_{outer}$. %JMC/SL - the term fierce sounds silly and non-professional (sorry!) I took out the place where you defined it, and we only use it in two other places. No need to invent a definition only to use it 2-3 times...can you swap in different, more descriptive words? / Again, I am trying to be consistent with what Jason and Haocheng used. I agree that 'fierce' is not the best word but if it is different from what J/H did, then there are inconsistencies between the publications...
We set the minimum threshold \emph{\textbf{g}} to be 730, found through experimentation to provide a good balance between performance and safety. %JMC/SL - This description is not great. The last sentence's use of g doesn't make sense with the equation...is g a threshold or the sum? Please try to clean up this description. / Think this clarified things

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=gaitspace.eps, width=8cm}
% \caption[ ]{A map of possible leg positions. Manually placed position cropping systems limited possible leg positions to avoid servo shutdowns and the robot hitting itself. The sum of the inner and outer servo positions cannot drop below 730, the smart cropping border, marked by the red line. Figure taken from Shen et al. \cite{haocheng}.}
% \end{center}
% \end{figure}
\figp{smartCrop}{.6}{A map of possible leg positions. Manually placed position cropping systems limited possible leg positions to avoid servo shutdowns and the robot hitting itself. The sum of the inner and outer servo positions cannot drop below 730, the smart cropping border, marked by the red line. Figure taken from Shen et al. \cite{haocheng}.}

>>>>

The Dynamixel AX-12A and AX-18A servos have a safety feature called \emph{Alarm Shutdown} that would turn off the servo if the servo's current, range, temperature, or torque was too high \cite{robotis}. 
The ones most relevant to this study were shutdowns due to high torque and range. 
As has been reported previously with QuadraTot \cite{yos:clune}, many of the gaits that were tested on the QuadraTot forced the servos to exert too much torque or would command out of range positions and cause the servos to shutdown. In order to address this problem and avoid collisions with its own body, a motion cropping system was implemented on the QuadraTot. 
In addition to placing limits on the range of motions for the inner joints, outer joints, and the hip joint of [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree] respectively, an extra cropping system named the smart cropping system was adopted from Shen et al. \cite{haocheng} to further prevent servo shutdowns from leg positions where the combination of the two angles interact to produce an extreme amount of torque. %This smart cropping algorithm calculates a leg position by summing the servo position values from the inner and outer joints of a leg, \emph{\textbf{g}} (\emph{\textbf{g}} = $j_{inner}$ + $j_{outer}$) %JMC/SL - do they sum the position values, or the joint angles? Since all the previous extremes are expressed in angles, the writing makes me expect angles... / They do use the position values and I think that it would be best to be consistent with what Jason/Haocheng did. This is also the way g is calculated in the simulator and the python drivers as well and more accurately reflects what we actually did. 
%and modifying the servo positions from ``fierce gaits'', which usually have relatively low \emph{\textbf{g}}s, to remap them onto the boundary of a quasi-trangular area in a 2-D space defined by $j_{inner}$ and $j_{outer}$. %JMC/SL - the term fierce sounds silly and non-professional (sorry!) I took out the place where you defined it, and we only use it in two other places. No need to invent a definition only to use it 2-3 times...can you swap in different, more descriptive words? / Again, I am trying to be consistent with what Jason and Haocheng used. I agree that 'fierce' is not the best word but if it is different from what J/H did, then there are inconsistencies between the publications...
%We set the minimum threshold \emph{\textbf{g}} to be 730, found through experimentation to provide a good balance between performance and safety. %JMC/SL - This description is not great. The last sentence's use of g doesn't make sense with the equation...is g a threshold or the sum? Please try to clean up this description. / Think this clarified things

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=gaitspace.eps, width=8cm}
% \caption[ ]{A map of possible leg positions. Manually placed position cropping systems limited possible leg positions to avoid servo shutdowns and the robot hitting itself. The sum of the inner and outer servo positions cannot drop below 730, the smart cropping border, marked by the red line. Figure taken from Shen et al. \cite{haocheng}.}
% \end{center}
% \end{figure}
%\figp{smartCrop}{.6}{A map of possible leg positions. Manually placed position cropping systems limited possible leg positions to avoid servo shutdowns and the robot hitting itself. The sum of the inner and outer servo positions cannot drop below 730, the smart cropping border, marked by the red line. Figure taken from Shen et al. \cite{haocheng}.}

### 2012-09-12 paper_sean/simulator.tex
lots of cuts, down to 13 pages.

% !TEX root =main.tex
Simulations occurred in C++ software customized for the QuadraTot\footnote{Available for Linux and Windows at http://quadratot.yosinski.com} in the NVIDIA PhysX physics simulator, as described previously \cite{glette}. 
Specifically, the simulation captured the number and type of joints, mass, and rough shapes and lengths of the QuadraTot components. 
%The simulator was visualized by OpenGL library to allow the user to have a more intuitive grasp of what is going on. %JMC Unnecessary detail since this paper is not about the simulator. 
%The simulator visualization included a recording feature which allows the user to create videos of the robot in simulation.%JMC: Assumed
%The simulator also features a command line function which allows the users to input a gait file and run it on the simulator. %JMC: Assumed
%The simulator then outputs a file containing the X, Y, Z coordinates of the center of mass of the robot over time. %JMC: Assumed
%In simulation, the robot traverses in the XY plane with the Z axis representing the height off the ground. %JMC Unnecessary detail
In an attempt to more accurately model the robot behavior, we implemented the same cropping function in the simulator as in the actual robot. %JMC/SL - Do all results in this paper reflect this? / They will soon.
Although the simulator was originally written in Windows for the Windows operating platform we ported and ran it on Linux (Ubuntu 11.04) to seamlessly link the simulator and the already exising libraries together. 
The friction values used for this study was the same as that used in Glette et al. \cite{glette}. 
%
% and uses OpenGL and GLFW libraries for visualization. The simula- tor was originally written for the Windows operating system but was ported and run on Linux (Ubuntu 11.04) for this ex- periment. It features a commandline function which allows the user to specify a gait input file and extract an output file with the X, Y, Z coordinate information over time. The Eu- clidean distance travelled by the robot in the XY plane was used to calculate the speed of the robot in simulation and was used as the fitness value used for evolution.
%

>>>>

% !TEX root =main.tex
Simulations occurred in C++ software customized for the QuadraTot\footnote{Available for Linux and Windows at http://quadratot.yosinski.com} in the NVIDIA PhysX physics simulator, as described previously \cite{glette}. 
Specifically, the simulation captured the number and type of joints, mass, and rough shapes and lengths of the QuadraTot components. 
%The simulator was visualized by OpenGL library to allow the user to have a more intuitive grasp of what is going on. %JMC Unnecessary detail since this paper is not about the simulator. 
%The simulator visualization included a recording feature which allows the user to create videos of the robot in simulation.%JMC: Assumed
%The simulator also features a command line function which allows the users to input a gait file and run it on the simulator. %JMC: Assumed
%The simulator then outputs a file containing the X, Y, Z coordinates of the center of mass of the robot over time. %JMC: Assumed
%In simulation, the robot traverses in the XY plane with the Z axis representing the height off the ground. %JMC Unnecessary detail
We implemented the initial cropping function in the simulator as in the actual robot but not the smart cropping function because with the smart cropping function implemented in simulation, there was too much limitations in the movements which drastically limited the possible gaitspace HyperNEAT could explore. %JMC/SL - Do all results in this paper reflect this? / They will soon.
Although the simulator was originally written in Windows for the Windows operating platform we ported and ran it on Linux (Ubuntu 11.04) to seamlessly link the simulator and the already exising libraries together. 
The friction values used for this study was the same as that used in Glette et al. \cite{glette}. 
%
% and uses OpenGL and GLFW libraries for visualization. The simula- tor was originally written for the Windows operating system but was ported and run on Linux (Ubuntu 11.04) for this ex- periment. It features a commandline function which allows the user to specify a gait input file and extract an output file with the X, Y, Z coordinate information over time. The Eu- clidean distance travelled by the robot in the XY plane was used to calculate the speed of the robot in simulation and was used as the fitness value used for evolution.
%
